---
title: "GLM Analysis (Trial-Level) — MST"
author: "Abhishek Garimella"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    latex_engine: xelatex
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(performance)
library(tidyverse)
library(lme4)
library(lmerTest)     # Satterthwaite p-values for mixed models
library(emmeans)      # post-hoc contrasts (Tukey / Bonferroni)
library(parameters)   # model_parameters(): clean CI + p tables
library(effectsize)   # Cohen's d, eta-squared
library(performance)  # check_model() diagnostics
library(pwr)          # post-hoc power analysis
```

---

## 1. Configuration

```{r config}
GROUP_FOLDERS <- list(
  "Task-Only"  = "./MST_Data/task_only/task_only_data",
  "Both"       = "./MST_Data/Both_item_task/both_data",
  "Item-Only"  = "./MST_Data/item_only/item_only_data"
)

BIN_FILE <- "./MST_Data/Both_item_task/Set6 bins.txt"
```

---

## 2. Data Loading & Parsing

```{r load-data}
classify_image <- function(path) {
  path <- gsub("\\\\", "/", path)
  bn   <- tolower(basename(path))
  if (grepl("foil", tolower(path))) return("Foil")
  if (grepl("a\\.jpg$", bn))        return("Target")
  if (grepl("b\\.jpg$", bn))        return("Lure")
  return(NA_character_)
}

parse_task <- function(filepath, group) {
  df   <- read_csv(filepath, show_col_types = FALSE)
  main <- if ("trials.thisN" %in% names(df)) filter(df, !is.na(`trials.thisN`)) else df
  if (nrow(main) == 0) return(NULL)
  pid  <- substr(basename(filepath), 1, 5)
  acc  <- if ("encoding_task_accuracy" %in% names(main))
            mean(main$encoding_task_accuracy, na.rm = TRUE) else NA_real_
  tibble(
    participant = pid,
    subject_id  = paste0(group, "_", pid),   # ← unique ID
    enc_accuracy = acc,
    group = group
  )
}

parse_test <- function(filepath, group) {
  df     <- read_csv(filepath, show_col_types = FALSE)
  trials <- if ("trials.thisN" %in% names(df)) filter(df, !is.na(`trials.thisN`)) else df
  needed <- c("image_path", "position_of_stimuli",
              "trials.key_resp_3.keys", "trials.key_resp_3.rt")
  if (!all(needed %in% names(trials)) || nrow(trials) == 0) return(NULL)
  pid <- substr(basename(filepath), 1, 5)
  trials %>%
    select(all_of(needed)) %>%
    rename(image_path = 1, position = 2, response = 3, rt = 4) %>%
    mutate(
      participant = pid,
      subject_id  = paste0(group, "_", pid),   # ← unique ID
      group       = group,
      img_type    = map_chr(image_path, classify_image),
      position    = tolower(replace_na(as.character(position), "none")),
      image_id    = basename(image_path)
    )
}

all_task <- list()
all_test <- list()

for (grp in names(GROUP_FOLDERS)) {
  folder     <- GROUP_FOLDERS[[grp]]
  files      <- sort(list.files(folder, pattern = "\\.csv$", full.names = TRUE))
  task_files <- files[grepl("_task_", basename(files))]
  test_files <- files[grepl("_test_", basename(files))]
  task_ids   <- setNames(task_files, substr(basename(task_files), 1, 5))
  test_ids   <- setNames(test_files, substr(basename(test_files), 1, 5))
  valid_ids  <- intersect(names(task_ids), names(test_ids))
  cat(grp, ": paired participants =", length(valid_ids), "\n")
  for (pid in valid_ids) {
    t  <- try(parse_task(task_ids[pid], grp), silent = TRUE)
    te <- try(parse_test(test_ids[pid], grp), silent = TRUE)
    if (!inherits(t,  "try-error") && !is.null(t))  all_task[[length(all_task)+1]]  <- t
    if (!inherits(te, "try-error") && !is.null(te)) all_test[[length(all_test)+1]]  <- te
  }
}

task_info   <- bind_rows(all_task)
test_trials <- bind_rows(all_test)

# now correctly reports 158
cat("Task participants parsed:", n_distinct(task_info$subject_id), "\n")
cat("Test participants parsed:", n_distinct(test_trials$subject_id), "\n")
```

---

## 3. Preprocessing, Exclusions & Feature Engineering

```{r preprocess}
test_trials <- test_trials %>%
  inner_join(task_info, by = c("participant", "group"))

# ── FIX: create unique subject ID to prevent PID collision across groups ──
test_trials <- test_trials %>%
  mutate(subject_id = paste0(group, "_", participant))

before <- n_distinct(test_trials$subject_id)

# --- Exclusion 1: low encoding accuracy (uncomment to enable) ---
# valid_sids <- test_trials %>%
#   filter(enc_accuracy >= 0.60) %>%
#   pull(subject_id) %>% unique()
# test_trials <- filter(test_trials, subject_id %in% valid_sids)

# --- Exclusion 2: high foil false-alarm rate > 0.50 (uncomment to enable) ---
# foil_far <- test_trials %>%
#   filter(img_type == "Foil") %>%
#   group_by(subject_id) %>%
#   summarise(far = mean(tolower(response) == "o", na.rm = TRUE))
# valid_sids2 <- foil_far %>% filter(far <= 0.50) %>% pull(subject_id)
# test_trials <- filter(test_trials, subject_id %in% valid_sids2)

after <- n_distinct(test_trials$subject_id)
cat("Excluded", before - after, "participants | Remaining:", after, "\n")
# should now print 158 instead of 72

# Binary outcomes
test_trials <- test_trials %>%
  mutate(
    resp_old     = as.integer(tolower(response) == "o"),
    resp_similar = as.integer(tolower(response) == "s"),
    pos_post     = as.integer(position == "post"),
    pos_pre      = as.integer(position == "pre"),
    grp_task     = as.integer(group == "Task-Only"),
    grp_both     = as.integer(group == "Both"),
    Position     = factor(position, levels = c("mid", "pre", "post")),
    Condition    = factor(group,    levels = c("Item-Only", "Task-Only", "Both")),
    subject_id   = factor(subject_id),   # factor for random effects
    participant  = factor(participant),
    image_id     = factor(image_id)
  )

# Lure similarity bins (1–5) from external mapping file
test_trials <- test_trials %>%
  mutate(
    stim_num    = as.numeric(str_extract(gsub("\\\\", "/", image_path),
                                         "(\\d+)[ab]\\.jpg", group = 1)),
    stim_idx120 = ((stim_num - 1) %% 120) + 1
  )

if (file.exists(BIN_FILE)) {
  bin_lookup  <- read_table(BIN_FILE, col_names = c("stim_idx120", "bin"), col_types = "ii")
  test_trials <- left_join(test_trials, bin_lookup, by = "stim_idx120")
} else {
  warning(paste("Bin file not found:", BIN_FILE, "— using stim_idx120 as fallback."))
  test_trials$bin <- test_trials$stim_idx120
}

test_trials$bin <- as.integer(test_trials$bin)
cat("Rows available for modeling:", nrow(test_trials), "\n")
cat("Non-null bin rows:",           sum(!is.na(test_trials$bin)), "\n")
cat("Unique subjects:",             n_distinct(test_trials$subject_id), "\n")  # verify = 158
```

---

## 4. GLMM — REC (Binomial, logit link)

Predicts trial-level `resp_old` (0/1): Position (post/pre vs mid) × Group,
with random intercepts for participant and image.

```{r glmm-rec}
rec_data <- test_trials %>%
  filter(img_type %in% c("Target", "Foil")) %>%
  drop_na(resp_old)

# Formula mirrors Python: pos_post*grp_task + pos_post*grp_both + pos_pre*grp_task + pos_pre*grp_both
glmm_rec <- glmer(
  resp_old ~ pos_post * grp_task + pos_post * grp_both +
             pos_pre  * grp_task + pos_pre  * grp_both +
             (1 | subject_id) + (1 | image_id),
  data    = rec_data,
  family  = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(glmm_rec)
```

```{r glmm-rec-table}
# Full fixed-effects table: beta (log-OR), SE, z, p, 95% CI (Wald), OR
cat("\n--- Fixed Effects: log-OR scale ---\n")
model_parameters(glmm_rec, exponentiate = FALSE, ci = 0.95) %>% print()

cat("\n--- Fixed Effects: Odds Ratio scale ---\n")
model_parameters(glmm_rec, exponentiate = TRUE,  ci = 0.95) %>% print()
```
```{r}
check_overdispersion(glmm_rec)
check_collinearity(glmm_rec)
```

```{r}
re_participant_rec <- ranef(glmm_rec)$subject_id[, 1]
re_image_rec       <- ranef(glmm_rec)$image_id[, 1]

# Q-Q plots for both
par(mfrow = c(1, 2))
qqnorm(re_participant_rec, main = "REC: Participant Random Intercepts")
qqline(re_participant_rec, col = "red")

qqnorm(re_image_rec, main = "REC: Image Random Intercepts")
qqline(re_image_rec, col = "red")

# Histograms
p_re1 <- ggplot(data.frame(re = re_participant_rec), aes(x = re)) +
  geom_histogram(bins = 15, fill = "#4C8BE0", colour = "white") +
  labs(title = "REC: Participant Random Intercepts", x = "b_i") +
  theme_minimal()

p_re2 <- ggplot(data.frame(re = re_image_rec), aes(x = re)) +
  geom_histogram(bins = 20, fill = "#E04C4C", colour = "white") +
  labs(title = "REC: Image Random Intercepts", x = "b_j") +
  theme_minimal()
```
```{r}
library(lme4)

# Pearson residuals
rec_data$pearson_resid <- residuals(glmm_rec, type = "pearson")

# Check if residuals correlate with trial order within participants
rec_data %>%
  group_by(subject_id) %>%
  mutate(trial_order = row_number()) %>%
  summarise(autocor = cor(pearson_resid, trial_order, use = "complete.obs")) %>%
  summarise(mean_autocor = mean(autocor, na.rm = TRUE))
```
```{r}
rec_data$obs_id <- seq_len(nrow(rec_data))

glmm_rec_olre <- glmer(
  resp_old ~ pos_post * grp_task + pos_post * grp_both +
             pos_pre  * grp_task + pos_pre  * grp_both +
             (1 | subject_id) + (1 | image_id) + (1 | obs_id),
  data    = rec_data,
  family  = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

# Compare with original
anova(glmm_rec, glmm_rec_olre)
```

```{r save-glmm-rec}
# Save REC fixed effects with requested columns
rec_coef <- as.data.frame(summary(glmm_rec)$coefficients)
rec_coef$term <- rownames(rec_coef)
rownames(rec_coef) <- NULL

rec_ci <- as.data.frame(confint(glmm_rec, method = "Wald", level = 0.95))
rec_ci$term <- rownames(rec_ci)
rownames(rec_ci) <- NULL
rec_ci <- rec_ci %>%
  filter(term %in% rec_coef$term) %>%
  select(term, `2.5 %`, `97.5 %`)

rec_out <- rec_coef %>%
  left_join(rec_ci, by = "term") %>%
  transmute(
    term,
    `log odds` = Estimate,
    `SE_log odds` = `Std. Error`,
    `95%CI_logodds` = sprintf("[%.6f, %.6f]", `2.5 %`, `97.5 %`),
    `odds ratio` = exp(Estimate),
    `SE_OR` = exp(Estimate) * `Std. Error`,
    `95%CI_oR` = sprintf("[%.6f, %.6f]", exp(`2.5 %`), exp(`97.5 %`)),
    z = `z value`,
    p = `Pr(>|z|)`
  )

write_csv(rec_out, "./GLMM_REC.csv")
cat("Saved: GLMM_REC.csv\n")

# Save REC random effects: Groups, Name, Variance, Std.Dev
rec_randeff <- as.data.frame(VarCorr(glmm_rec)) %>%
  transmute(
    Groups = grp,
    Name = if_else(is.na(var1), "(Intercept)", var1),
    Variance = vcov,
    `Std.Dev` = sdcor
  )

write_csv(rec_randeff, "./GLMM_RandEff_REC.csv")
cat("Saved: GLMM_RandEff_REC.csv\n")
```

---

## 5. Post-hoc Contrasts — REC (Tukey's HSD + Bonferroni)

Runs because REC gave a significant result in the Python notebook.

```{r posthoc-rec}
# EMMs on the dummy-coded model — refit using factor columns for cleaner emmeans output
glmm_rec_fac <- glmer(
  resp_old ~ Position * Condition + (1 | subject_id) + (1 | image_id),
  data    = rec_data,
  family  = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

emm_rec <- emmeans(glmm_rec_fac, ~ Position | Condition)

cat("\n--- Tukey HSD: pairwise Position contrasts within each Condition ---\n")
pairs(emm_rec, adjust = "tukey") %>% as_tibble() %>% arrange(Condition, p.value) %>% print()

cat("\n--- Bonferroni: Post vs Mid contrast per Condition ---\n")
contrast(emm_rec, method = list("Post-Mid" = c(-1, 0, 1)), adjust = "bonferroni") %>%
  as_tibble() %>% print()
```

```{r save-posthoc-rec}
# Save REC post-hoc contrasts with requested columns
posthoc_tukey <- pairs(emm_rec, adjust = "tukey") %>%
  as_tibble() %>%
  mutate(
    correction = "tukey",
    `cohen's d` = estimate / SE
  ) %>%
  transmute(
    correction,
    contrast,
    condition = Condition,
    estimate,
    SE,
    df,
    z.ratio,
    p.value,
    `cohen's d`
  )

posthoc_bonf <- contrast(emm_rec, method = list("Post-Mid" = c(-1, 0, 1)), adjust = "bonferroni") %>%
  as_tibble() %>%
  mutate(
    correction = "bonferroni",
    `cohen's d` = estimate / SE
  ) %>%
  transmute(
    correction,
    contrast,
    condition = Condition,
    estimate,
    SE,
    df,
    z.ratio,
    p.value,
    `cohen's d`
  )

posthoc_out <- bind_rows(posthoc_tukey, posthoc_bonf)
write_csv(posthoc_out, "./GLMM_PostHoc_REC.csv")
cat("Saved: GLMM_PostHoc_REC.csv\n")
```

---

## 6. GLMM — LDI (Binomial, logit link, with Similarity Bins)

Predicts trial-level `resp_similar` (0/1): pre-boundary × group interactions,
similarity bin main effect, and Position × Bin interaction.

```{r glmm-ldi}
ldi_data <- test_trials %>%
  filter(img_type %in% c("Lure", "Foil")) %>%
  drop_na(resp_similar, bin)

# Formula mirrors Python exactly
glmm_ldi <- glmer(
  resp_similar ~ pos_pre * grp_task + pos_pre * grp_both +
                 factor(bin) + pos_pre:factor(bin) +
                 (1 | subject_id) + (1 | image_id),
  data    = ldi_data,
  family  = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

summary(glmm_ldi)
```

```{r glmm-ldi-table}
cat("\n--- Fixed Effects: log-OR scale ---\n")
model_parameters(glmm_ldi, exponentiate = FALSE, ci = 0.95) %>% print()

cat("\n--- Fixed Effects: Odds Ratio scale ---\n")
model_parameters(glmm_ldi, exponentiate = TRUE,  ci = 0.95) %>% print()
```

```{r}
check_overdispersion(glmm_ldi)
check_collinearity(glmm_ldi)
```
```{r}
re_participant_ldi <- ranef(glmm_ldi)$subject_id[, 1]
re_image_ldi       <- ranef(glmm_ldi)$image_id[, 1]

# Q-Q plots for both
par(mfrow = c(1, 2))
qqnorm(re_participant_ldi, main = "LDI: Participant Random Intercepts")
qqline(re_participant_ldi, col = "red")

qqnorm(re_image_ldi, main = "LDI: Image Random Intercepts")
qqline(re_image_ldi, col = "red")

# Histograms
p_re1 <- ggplot(data.frame(re = re_participant_ldi), aes(x = re)) +
  geom_histogram(bins = 15, fill = "#4C8BE0", colour = "white") +
  labs(title = "LDI: Participant Random Intercepts", x = "b_i") +
  theme_minimal()

p_re2 <- ggplot(data.frame(re = re_image_ldi), aes(x = re)) +
  geom_histogram(bins = 20, fill = "#E04C4C", colour = "white") +
  labs(title = "LDI: Image Random Intercepts", x = "b_j") +
  theme_minimal()
```


```{r}
library(lme4)

# Pearson residuals
ldi_data$pearson_resid <- residuals(glmm_ldi, type = "pearson")

# Check if residuals correlate with trial order within participants
ldi_data %>%
  group_by(subject_id) %>%
  mutate(trial_order = row_number()) %>%
  summarise(autocor = cor(pearson_resid, trial_order, use = "complete.obs")) %>%
  summarise(mean_autocor = mean(autocor, na.rm = TRUE))
```
```{r}
ldi_data$obs_id <- seq_len(nrow(ldi_data))
glmm_ldi_olre <- glmer(
  resp_similar ~ pos_pre * grp_task + pos_pre * grp_both +
                 factor(bin) + pos_pre:factor(bin) +
                 (1 | subject_id) + (1 | image_id) + (1 | obs_id),
  data    = ldi_data,
  family  = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)
# Compare with original
anova(glmm_ldi, glmm_ldi_olre)
```

```{r save-glmm-ldi}
# Save LDI fixed effects with requested columns
ldi_coef <- as.data.frame(summary(glmm_ldi)$coefficients)
ldi_coef$term <- rownames(ldi_coef)
rownames(ldi_coef) <- NULL

ldi_ci <- as.data.frame(confint(glmm_ldi, method = "Wald", level = 0.95))
ldi_ci$term <- rownames(ldi_ci)
rownames(ldi_ci) <- NULL
ldi_ci <- ldi_ci %>%
  filter(term %in% ldi_coef$term) %>%
  select(term, `2.5 %`, `97.5 %`)

ldi_out <- ldi_coef %>%
  left_join(ldi_ci, by = "term") %>%
  transmute(
    term,
    `log odds` = Estimate,
    `SE_log odds` = `Std. Error`,
    `95%CI_logodds` = sprintf("[%.6f, %.6f]", `2.5 %`, `97.5 %`),
    `odds ratio` = exp(Estimate),
    `SE_OR` = exp(Estimate) * `Std. Error`,
    `95%CI_oR` = sprintf("[%.6f, %.6f]", exp(`2.5 %`), exp(`97.5 %`)),
    z = `z value`,
    p = `Pr(>|z|)`
  )

write_csv(ldi_out, "./GLMM_LDI.csv")
cat("Saved: GLMM_LDI.csv\n")
```

```{r glmm-ldi-bin-interaction}
# Inspect Position × Bin interaction: does pre-boundary effect vary by bin?
glmm_ldi_fac <- glmer(
  resp_similar ~ pos_pre * Condition + factor(bin) + pos_pre:factor(bin) +
                 (1 | subject_id) + (1 | image_id),
  data    = ldi_data,
  family  = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

emm_ldi_bin <- emmeans(glmm_ldi_fac, ~ pos_pre | bin)
cat("\n--- Position (Pre vs Mid) contrast at each similarity bin ---\n")
pairs(emm_ldi_bin, adjust = "tukey") %>% as_tibble() %>% arrange(bin) %>% print()
```


---

## 7. GLM — RT Slowing → REC Deficit (Participant-Level Regression)

Replicates the OLS regression from cell 5 of the Python notebook.
```{r rt-delta}
get_rt_delta <- function(filepath) {
  df <- read_csv(filepath, show_col_types = FALSE,
                 na = c("", "NA", "None", "nan", "NaN"))

  main <- if ("trials.thisN" %in% names(df))
            filter(df, !is.na(`trials.thisN`)) else df
  if (nrow(main) == 0) return(NULL)

  pid <- substr(basename(filepath), 1, 5)

  main <- main %>%
    mutate(
      pos_idx = as.integer(`trials.thisN`) %% 7,
      pos     = case_when(
        pos_idx == 0 ~ "post",
        pos_idx == 6 ~ "pre",
        TRUE         ~ "mid"
      )
    )

  rt_candidates <- intersect(
    c("trials.key_resp_8.rt", "trials.key_resp_9.rt"),
    names(main)
  )
  if (length(rt_candidates) == 0) return(NULL)

  # Keep only columns with at least one valid numeric RT
  rt_candidates <- rt_candidates[sapply(rt_candidates, function(x) {
    vals <- suppressWarnings(as.numeric(main[[x]]))
    any(!is.na(vals))
  })]
  if (length(rt_candidates) == 0) return(NULL)

  rt_col <- rt_candidates[1]

  rt_by_pos <- main %>%
    mutate(rt_value = suppressWarnings(as.numeric(.data[[rt_col]]))) %>%
    group_by(pos) %>%
    summarise(med_rt = median(rt_value, na.rm = TRUE), .groups = "drop")

  mid_rt  <- rt_by_pos$med_rt[rt_by_pos$pos == "mid"]
  post_rt <- rt_by_pos$med_rt[rt_by_pos$pos == "post"]

  if (length(mid_rt) == 0 || length(post_rt) == 0 ||
      is.na(mid_rt) || is.na(post_rt)) return(NULL)

  tibble(participant = pid, rt_delta_post = post_rt - mid_rt)
}

task_files_rt <- unlist(lapply(GROUP_FOLDERS, function(f)
  sort(list.files(f, pattern = "_MST_task_.*\\.csv$", full.names = TRUE))))

rt_data <- bind_rows(lapply(task_files_rt, get_rt_delta))
```

```{r rec-agg}
foil_far_df <- test_trials %>%
  filter(img_type == "Foil") %>%
  group_by(participant) %>%
  summarise(foil_far = mean(resp_old, na.rm = TRUE))

rec_agg <- test_trials %>%
  filter(img_type == "Target") %>%
  group_by(participant, position) %>%
  summarise(hit_rate = mean(resp_old, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = position, values_from = hit_rate) %>%
  left_join(foil_far_df, by = "participant") %>%
  mutate(
    REC_post       = post - foil_far,
    REC_mid        = mid  - foil_far,
    REC_delta_post = REC_post - REC_mid
  ) %>%
  left_join(distinct(task_info, participant, group), by = "participant") %>%
  mutate(subject_id = paste0(group, "_", participant)) %>%
  select(participant, group, subject_id, REC_delta_post)
```

```{r glm-rt-rec}
rt_data_grp <- rt_data %>%
  left_join(distinct(task_info, participant, group), by = "participant") %>%
  mutate(subject_id = paste0(group, "_", participant))

reg_data <- rt_data_grp %>%
  inner_join(rec_agg, by = "subject_id") %>%
  select(subject_id,
         participant = participant.x,
         group       = group.x,
         rt_delta_post,
         REC_delta_post) %>%
  drop_na(rt_delta_post, REC_delta_post) %>%
  # ── FIX: collapse duplicate rows by averaging ──────────────────────────
  group_by(subject_id, participant, group) %>%
  summarise(
    rt_delta_post  = mean(rt_delta_post,  na.rm = TRUE),
    REC_delta_post = mean(REC_delta_post, na.rm = TRUE),
    .groups = "drop"
  )

# ── Verify: should now be exactly 1 row per subject ───────────────────────
cat("N for regression:", nrow(reg_data), "\n")

glm_rt <- lm(REC_delta_post ~ rt_delta_post + factor(group), data = reg_data)
summary(glm_rt)
```


```{r glm-rt-table}
# Full table with beta, SE, t, p, 95% CI
cat("\n--- Fixed Effects with 95% CIs ---\n")
model_parameters(glm_rt, ci = 0.95) %>% print()
```

```{r glm-rt-correlation}
# Overall Pearson r (mirrors Python output)
ct <- cor.test(reg_data$rt_delta_post, reg_data$REC_delta_post)
cat(sprintf("\nOverall Pearson r (RT slowing vs REC deficit): r=%.3f, p=%.3f\n", ct$estimate, ct$p.value))
cat("95% CI: [", round(ct$conf.int, 3), ",", round(ct$conf.int, 3), "]\n")[2][1]
```

```{r glm-assumption1-linearity}
library(tidyverse); library(ggplot2); library(car); library(lmtest); library(sandwich)

# Linearity: scatter + LOESS vs linear
ggplot(reg_data, aes(x = rt_delta_post, y = REC_delta_post)) +
  geom_point(alpha = 0.5, colour = "#4C8BE0") +
  geom_smooth(method = "lm",    se = TRUE,  colour = "#E04C4C",
              linetype = "solid",  linewidth = 1) +
  geom_smooth(method = "loess", se = FALSE, colour = "#4CBE8B",
              linetype = "dashed", linewidth = 1) +
  labs(title    = "Assumption 1: Linearity",
       subtitle = "Red = linear fit | Green dashed = LOESS\nShould overlap if relationship is linear",
       x = "Δ RT (Post − Mid, s)", y = "Δ REC (Post − Mid, corrected)") +
  theme_minimal(base_size = 13)
```

```{r glm-assumption1b-residuals-vs-fitted}
# Residuals vs Fitted — formal linearity check
ggplot(data.frame(fitted    = fitted(glm_rt),
                  residuals = residuals(glm_rt)),
       aes(x = fitted, y = residuals)) +
  geom_point(alpha = 0.5, colour = "#4C8BE0") +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "#E04C4C") +
  geom_smooth(method = "loess", se = FALSE, colour = "#4CBE8B", linewidth = 1) +
  labs(title    = "Assumption 1b: Residuals vs Fitted",
       subtitle = "LOESS should be flat around 0 — any curve = nonlinearity",
       x = "Fitted Values", y = "Residuals") +
  theme_minimal(base_size = 13)
```

```{r glm-assumption2-independence}
# Independence: Durbin-Watson
dw <- dwtest(glm_rt)
print(dw)
cat(sprintf("DW statistic = %.3f (ideal ≈ 2.0)\n", dw$statistic))
cat(if (dw$p.value > 0.05) "No autocorrelation\n" else "Autocorrelation detected\n")
```

```{r glm-assumption3a-qq}
# Normality: Q-Q plot

ggplot(data.frame(resid = residuals(glm_rt)), aes(sample = resid)) +
  stat_qq(colour = "#7B4CE0", alpha = 0.7) +
  stat_qq_line(colour = "#E04C4C", linewidth = 1) +
  labs(title    = "Assumption 3: Residual Normality (Q-Q)",
       subtitle = "Points should hug the red diagonal",
       x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_minimal(base_size = 13)
```

```{r glm-assumption3b-histogram}
# Normality: Histogram
ggplot(data.frame(resid = residuals(glm_rt)), aes(x = resid)) +
  geom_histogram(aes(y = after_stat(density)), bins = 20,
                 fill = "#7B4CE0", colour = "white", alpha = 0.8) +
  geom_density(colour = "#E04C4C", linewidth = 1) +
  stat_function(fun  = dnorm,
                args = list(mean = mean(residuals(glm_rt)),
                            sd   = sd(residuals(glm_rt))),
                linetype = "dashed", colour = "black", linewidth = 0.9) +
  labs(title    = "Assumption 3: Residual Distribution",
       subtitle = "Dashed = Normal reference",
       x = "Residual", y = "Density") +
  theme_minimal(base_size = 13)
```

```{r glm-assumption4-homoscedasticity}
# Homoscedasticity: Breusch-Pagan + NCV
bp <- bptest(glm_rt)
print(bp)
cat(if (bp$p.value > 0.05) "Homoscedastic\n" else "Heteroscedasticity — consider robust SEs\n")

cat("\n--- NCV Test ---\n")
ncvTest(glm_rt)
```

```{r glm-assumption5-multicollinearity}
# Multicollinearity: VIF
vif_vals <- vif(glm_rt)
print(vif_vals)
cat(if (all(vif_vals < 5)) "No multicollinearity (all VIF < 5)\n" else "High VIF\n")

# RT delta means by group — check predictor-group independence
reg_data %>%
  group_by(group) %>%
  summarise(mean_rt = mean(rt_delta_post, na.rm = TRUE),
            sd_rt   = sd(rt_delta_post,   na.rm = TRUE),
            n = n())
```

```{r glm-assumption6-cooks}
# Step 1: Check actual Cook's D values — are any > 0.5?
cooksd <- cooks.distance(glm_rt)
cat("Max Cook's D:", round(max(cooksd), 4), "\n")
cat("Any Cook's D > 0.5?", any(cooksd > 0.5), "\n")
cat("Any Cook's D > 1.0?", any(cooksd > 1.0), "\n")

# Step 2: does removing them change conclusion?
reg_data_no_inf <- reg_data %>%
  mutate(cooksd = cooksd) %>%
  filter(cooksd <= 4/nrow(reg_data))

glm_rt_sens <- lm(REC_delta_post ~ rt_delta_post + factor(group),
                  data = reg_data_no_inf)

cat("\n--- Original result ---\n")
cat(sprintf("rt_delta_post: β=%.4f, p=%.4f\n",
    coef(glm_rt)["rt_delta_post"],
    summary(glm_rt)$coef["rt_delta_post", 4]))

cat("\n--- Without influential points ---\n")
cat(sprintf("rt_delta_post: β=%.4f, p=%.4f\n",
    coef(glm_rt_sens)["rt_delta_post"],
    summary(glm_rt_sens)$coef["rt_delta_post", 4]))

cat("\nConclusion stable?",
    if (summary(glm_rt_sens)$coef["rt_delta_post", 4] > 0.05)
    "YES null result holds" else "NO — influential points matter")
```

```{r save-rt-rec}
# Save RT -> REC regression summary with requested columns
rt_coef <- as.data.frame(summary(glm_rt)$coefficients)
rt_coef$Parameter <- rownames(rt_coef)
rownames(rt_coef) <- NULL

rt_ci <- as.data.frame(confint(glm_rt, level = 0.95))
rt_ci$Parameter <- rownames(rt_ci)
rownames(rt_ci) <- NULL
rt_ci <- rt_ci %>% rename(ci_low = `2.5 %`, ci_high = `97.5 %`)

rt_out <- rt_coef %>%
  left_join(rt_ci, by = "Parameter") %>%
  transmute(
    Parameter,
    Coefficient = Estimate,
    SE = `Std. Error`,
    `95% CI` = sprintf("[%.6f, %.6f]", ci_low, ci_high),
    `t(338)` = `t value`,
    p = `Pr(>|t|)`
  )

rt_corr_row <- tibble(
  Parameter = sprintf("Overall Pearson r (RT slowing vs REC deficit): r=%.3f, p=%.2f", unname(ct$estimate), ct$p.value),
  Coefficient = unname(ct$estimate),
  SE = NA_real_,
  `95% CI` = sprintf("[%.3f, %.3f]", ct$conf.int[1], ct$conf.int[2]),
  `t(338)` = NA_real_,
  p = ct$p.value
)

rt_out <- bind_rows(rt_out, rt_corr_row)
write_csv(rt_out, "./RT_REC.csv")
cat("Saved: RT_REC.csv\n")
```


---

## 8. Effect Sizes

```{r effect-sizes}
# Eta-squared for RT → REC GLM
cat("\n--- Eta-squared (partial) for RT regression ---\n")
eta_squared(glm_rt, partial = TRUE) %>% print()

# Cohen's d for REC post-hoc contrasts
cat("\n--- Cohen's d for REC Position contrasts ---\n")
eff_rec <- pairs(emm_rec, adjust = "none") %>%
  as_tibble() %>%
  mutate(cohens_d = estimate / SE)
print(eff_rec)
```

---

## 9. Post-hoc Power Analysis

```{r power}
# Minimum detectable effect at N=51 (smallest group: Both), alpha=0.05, power=0.80
pwr.t.test(
  n           = 51,
  sig.level   = 0.05,
  power       = 0.80,
  type        = "one.sample",
  alternative = "two.sided"
)
# The returned d is the smallest effect your study was powered to detect
```
